Thermodynamic Structure of the Solar Corona: Tomographic Reconstructions and MHD Modeling

### 2020-03-10 response_to_review_SOLA-D-20-00015_v1.txt

Dear Editor,

Thank you very much for your email with the report on the first version of our manuscript ID SOLA-D-20-00015, titled "Thermodynamic Structure of the Solar Corona: Tomographic Reconstructions and MHD Modeling".

We deeply acknowledge the reviewer, whose report reflects her/his very careful and interested read. We found all queries very relevant, with useful and constructive criticisms and suggestions. Below, we copy/paste the full review. Following each query (Q) we include a corresponding answer (A), commenting the changes in the manuscript, in which all edited/new text is highlighted in red color font.

We addressed all points, as a result of which we believe the manuscript improved in both content and clarity. In particular, a major improvement in this new version of the draft relates to Q5 below.

@@

------------------------------beginning of review and response---------------------------
COMMENTS FOR THE AUTHOR:

Review of Thermodynamic Structure of the Solar corona: tomographic reconstructions and MHD modeling

This is a well-structured and clearly-written paper, presenting a detailed comparison of tomographic reconstructions of the low solar corona (which gives density and temperature), and a global MHD model. The work is important and valuable to the field, and is suitable for publication in Solar Physics. There are no major scientific concerns. The paper should be ready for publication if the following, mostly minor, concerns are addressed:

Medium concerns

-----------------------------------------------------------------------------------------
Q1. All of the multi-panel figures that show longitude-latitude maps (e.g. figure 1, 2, 3 etc) are too small, it is very difficult to see any details. There is a lot of white space between the panels that can be removed, and this would be one way to instantly make the maps larger.

@@

-----------------------------------------------------------------------------------------
Q2. Figure 3 and 11, bottom panels. I find it difficult to distinguish between the red and magenta colours. Can you replace magenta with a different colour, maybe green?

A2. We agree. Actually this affects all following Figures: 3, 4, 11, and 14. In those, we applied the following changes: Red->Orange and Magenta->Red. These emulate our previous choice of colors in Nuevo et al. (2013) and Lloveras et al. (2017). We like to keep all "up" structures in a "warm" palette. We edited text accordingly all along the manuscript. We believe the colors are now clearly distinguishable in all figures.

-----------------------------------------------------------------------------------------
Q3. Page 14, paragraph starting at line 34. You discuss the latitudinal distribution of short and long loops, and state that the small legs mainly populate the equatorial latitudes, and longer legs at higher latitudes. But this is a skewed result - their spatial positioning was a criteria for their categorisation, so this should not be presented as a scientific result.

A3. Several aspects are involved here, the discussion in the paragraph you indicate strongly relates on that in the final paragraph of Section 2.3. In reading them we find that these texts were confusing as they were, so we edited both. Actually, we do not say that *small* legs populate the equatorial latitudes and *large* ones the mid latitudes. We say that if one classifies the *small-closed* loops in DOWN and UP, the former peak near the equator and the latter peak near mid latitudes (at least for the deep minimum rotation CR-2082). The classification of legs in all four classes in Section 2.3 includes limits for the latitude of their footpoints, to distinguish different layers of the streamer belt (legs of type 0, I and II), as well as to separate field lines of the high latitude CHs (legs of type III). We included two new panels in Figure 3, showing frequency histograms of the latitude of the four types of legs at 1.105 Rsun. Within the range Abs(latitude) < 50 deg, and for the solar minimum target CR-2082, down-legs are peaked near the equator while up-legs peak at mid-latitudes. This is less obvious for CR-2208, although it is clear that down-legs peak in the core of the streamer belt. If one analyses Abs(latitude) > 50 deg it turns out that small-down legs are scarce and there is an extra population of small-up legs (not nearly as much as in the peaks seen at mid lats. Should have we not included the limit of Abs(latitude) < 50 deg for closed-small down and up legs (i.e., legs of type 0 and I), their histograms would still peak at around the same latitudes. This can be seen in Figure 7 of Nuevo et al. 92013), were they analyzed the small closed up and down loops for two rotations of solar minimum, without placing any limits in the latitude of the footpoints. As our main objective here is to be able to analyze progressively outer layers of the streamer belt, we include these limits in the latitude of the footpoints. We edited in full the final paragraph of Section 2.3, as well as the paragraph of Section 3.1 you point out (now split in two, shorter paragraphs). We trust the new graphs and discussions are clear now. Let us know.

@@

-----------------------------------------------------------------------------------------
Q4. Table 2 summarises the differences of the median values of the DEMT and MHD model. This is OK. However, there is an opportunity here to present a more detailed comparison, including in particular the mean (or median) absolute deviation (MAD) of values between DEMT and MHD. The MAD is a more stringent, voxel-by-voxel test of the comparison, which cannot be seen from the histogram comparisons (e.g. figure 12/13) nor the median values of Table 2.

A4. First of all we note that, as a result of the discussion in Q5 below, a few numbers in Table 2 were updated (those corresponding to type-III legs for CR-2208). Now, when selecting field lines for analysis of both the DEMT and the AWSoM models, the classification of types of legs introduced in section 2.3 is used for each model separately. As a result, as shown by comparing the middle panels of Figure 3 and the bottom panels of Figure 11, and as discussed in the corresponding texts of the manuscript, the selection of field lines for DEMT is not the same. In comparing those Figures each color indicates comparable regions between the two analysis. It is so that Figures 12 and 13, and Table 2, which show DEMT-versus-AWSoM comparison for each rotation, are not done line-by-line but rather region-by-region (say: orange, red and cyan). The suggestion of calculating the MAD would require selecting the same field lines for both models, which is beyond the scope of the present comparative analysis. Informing the MAD is a good idea, being a precise measure of how the comparison between both models spreads. We will certainly implement such an analysis in a next effort. In the present draft, we trust the current comparison (Figures 12 and 13, and Table 2) provide are pretty informative summary of how both models compare. Let us know what you think.

@@

-----------------------------------------------------------------------------------------
Q5. You discuss the less reliable results from AIA in coronal holes. The very recent paper by Saqri et al (Solar Physics 295, 2020) is relevant here, and should be referenced.

A5. Thanks for the reference, which is certainly relevant and now mentioned in the paper, as discussed below. We are not sure stray light contamination was actually the reason of the reported issue, which is now solved.

The discussion section of v1 of the draf include this sentence (now deleted): "The source of this distinct performance of DEMT when applied to EUVI or AIA data is related to the TRFs of the respective filter sets. We are currently implementing an improved LDEM determination algorithm to deal with this issue that will be used in a follow up article." We have successfully implemented the improved version of the code, as a result of which this problem has been solved. The issue was related to the algorithm that solves the optimization problem posed in each voxel to find the LDEM. When based on AIA data, the previous algorithm tended to converge to a non-optimal (local minimum of the objective function) in specific voxels located in the CHs. We addressed the problem by exploring multiple starting points when solving the optimization problem in cases where the initial solution was not good enough. In this draft, this improved algorithm affects results concerning type III legs for CR2208. As a result, the following Figures have been updated: 2, 3, 4, 5, 10, 13 and 14. You will notice all these Figures are just like the original ones, except the problem in the CHs of CR-2208 has been solved. Text throughout the manuscript was edited accordingly, in the following locations of v2 of the draft:
 - Red text in Pg 12 and Pg 13, and the red rext in Pg 15 informing characteristic values of Grad_r(Tm).
 - Numerical values concerning legs of type-III for CR-2208 in Tables 1 and 2.
 - Most of the third paragraph of the Conclusion section in V1 was devoted to this issue, so it was eliminated. The remaining text is now the end of the second paragraph of that section (highlighted in red color).

Concerning the effect of stray-light contamination of EUV images on DEMT products, our current data processing pipeline for EUVI data includes the deconvolution of the PSF, carefully determined by Shearer et al. (ApJL, 2012). We have not yet included such a procedure for AIA data, as we were not aware of a reliable determination of their PSF. Shearer et al. (2012) showed that the signal in the core of CHs may be strongly affected by stray light. Now, the effect in tomography turns out to be milder, due to temporal and spatial binning of the images. In the case of EUVI data, Lloveras et al. (2017) performed a careful analysis of effect of stray-light removal in the resulting DEMT products. PSF deconvolution basically increases the contrast. As a result, the off-limb radial brightness in EUV images becomes somewhat steeper, which has an impact in tomographic reconstructions. Specifically, Lloveras et al. (2017) -see the end of their section 3.6- found that the coronal base density increases a bit (5-15%) and the density scale height decreases a bit (5-10%), while the electron temperature remains largely unchanged. We understand that stray light correction is less serious for AIA data. Again, it may be strong in specific pixels of AIA images and yet the final effect in tomographic products may be mild. Still, we do want to explore the effect in DEMT of stray light removal in AIA data. It will be in particular interesting to see if stray-light removal helps the algorithm that finds de LDEM in voxels of the CHs. We acknowledge the reference on AIA data, and will investigate this in the future. We included a comment on this topic, as well as the reference by Saqri et.al., at the beginning of Section 2.1 (red text).

@@

-----------------------------------------------------------------------------------------
Q6. The abstract can be shortened somewhat, reducing the number of introductory sentences and with a little less emphasis on the method and future improvements, and more on the science results. In particular: the results of the energy balance analysis and the estimate of energy flux at the coronal base, which are important.

A6. We agree. We edited the Abstract. Let us know what you think.

-- FALTA ESTO --

-----------------------------------------------------------------------------------------
Q7. Page 5, paragraph beginning line 34. You briefly describe fitting the DEM to a Gaussian. I know this approach is covered in detail in Nuevo 2015, but I would like a little more detail here. You should mention that, in general, DEMs do not necessarily have a clean Gaussian profile and that more complicated inversions are needed (reference https://ui.adsabs.harvard.edu/abs/2012A%26A...539A.146H/abstract , https://ui.adsabs.harvard.edu/abs/2019SoPh..294..135M/abstract , https://ui.adsabs.harvard.edu/abs/2019SoPh..294..136P/abstract). Then explain briefly why DEMT leads to cleaner, non-multithermal profiles (due to the 3D spatial reconstruction of the line of sight). I assume that this is why you can use a straightforward Gaussian fit? Also efficiency?

A7. Thanks for this comment and the references. We were aware of those, discussing the first one in Lloveras et.al. (2017). Concerning the other two, we are in fact interested in using the SITES code in the determination of the LDEM at each voxel, and make comparisons with those provide by our current parametric model code. As for the reason of LDEMs being "cleaner" than DEMs, you are right. As the LDEM describes the temperature distribution of the plasma in a specific voxel it does not deal with different large scale structures, as it may be the case for the DEM describing the plasma along a full line-of-sight (LOS). In that sense one can expect as you say, cleaner, profiles than those observed by DEM studies. Even so, if one includes in the analysis data from the 335 A band, which adds two sensitivity points at ~ 1.0 MK and ~ 2.5 MK, one finds that a a bi-Gaussian LDEM model best predicts all the tomographic emissivities. That is the main conclusion of Nuevo et. al. (2015). The interpretation in that case is that the unresolved independent strands threading the voxel can have two different characteristic temperatures. In this sense, the plasma within the voxel can be multi-thermal. As for efficiency, the parametric technique is comparably efficient independently of the chosen model. We included a new paragraph (red text) between Equations (1) and (2), discussing this point and including the references above and others as well.

### 2020-03-10 lloveras_2020_v2.tex
Albert está trabajando en esto, no tocar todavía, gracias.

{As detailed in Section \ref{intro}, rotations CR-2082 and CR-2208 were selected to carry out DEMT reconstructions, based on data taken by the STEREO/EUVI-B and SDO/AIA instruments, respectively.} The EUVI and AIA data were prepared using the latest processing tools and calibration corrections provided by their teams through the SolarSoft package. For this work, we introduced two improvements in the implementation of the DEMT technique, as described next.

{While in all previous DEMT studies full-disk data was used to perform tomography, in this work we {opt} to only use off-limb data. In this way, the smallest scale and brighter coronal features seen on disk (most typically in the 171\AA\ band) are not included. This has two implications. Firstly, the fast dynamics that typically characterizes those structures is \edit{absent/reduced} from the data. Secondly, only half synodic rotation worth of data is needed to constrain the inversion problem for the whole coronal volume. As a result, {artifacts induced by coronal dynamics} are reduced compared to previous DEMT reconstructions.}

@@

@@

\edit{Legs of type down turn out to have a significant population only for closed-small field lines. Hence, we separate them into two complementary classes: legs of type 0 (down) and legs of type I (up). In the case of closed-large field lines, legs of type down are $\lesssim 15\%$ of the population. In the case of open field lines, legs of type down are $\lesssim 10\%$ of the population. Hence, the requirement of being up for legs of type II and III is included to select the vastly dominating population in each case. On the other hand, the inclusion of latitude limits for the footpoints in the classification of legs from type 0 through III is purposely set to study the streamer belt in three distinct, progressively outer layers, as well as to separate the field lines of the high latitude CHs (legs of type III). In Section \ref{resu}, the results of both the DEMT and AWSoM models in the four classes of legs are statistically analysed.}

@@

\begin{figure}[h!]
\begin{center}
\includegraphics[width=0.495\textwidth]{figs/map_Ne_CR2082_DEMT-EUVI_behind_H1-L3523_r3d_1025_Rsun.pdf}
\includegraphics[width=0.495\textwidth]{figs/map_Tm_CR2082_DEMT-EUVI_behind_H1-L3523_r3d_1025_Rsun.pdf}
\includegraphics[width=0.495\textwidth]{figs/map_Ne_CR2082_DEMT-EUVI_behind_H1-L3523_r3d_1065_Rsun.pdf}
\includegraphics[width=0.495\textwidth]{figs/map_Tm_CR2082_DEMT-EUVI_behind_H1-L3523_r3d_1065_Rsun.pdf}
\includegraphics[width=0.495\textwidth]{figs/map_Ne_CR2082_DEMT-EUVI_behind_H1-L3523_r3d_1105_Rsun.pdf}
\includegraphics[width=0.495\textwidth]{figs/map_Tm_CR2082_DEMT-EUVI_behind_H1-L3523_r3d_1105_Rsun.pdf}
\caption{Carrington maps of DEMT {products $\sqravgN$} (left panels) and $\Tm$ (right panels) for CR-2082. Top, middle and bottom panels show the results at three heliocentric heights, $1.025$, $1.065$ and $1.105\,\mrsun$ respectively. Black voxels correspond to non-reconstructed regions (see text in Section \ref{demt_res}) and thick-black curves indicate the open/closed boundaries. \comment{Por consistencia, creo que en los títulos de los CarrMaps de las Figuras 1 y 2 deberías usar la notación $\sqravgN$ y $\Tm$. Si esto lo hacés en LETITBE con el IDL 8.4, podrás usar los comandos de LaTeX dentro de las strings ``title", con lo cual no solo será simple sino que $\sqravgN$ no va a ocupar mucho más espacio y las chances son que salga todo lindo sin cambiar el tamaño de letra ni nada.}}
\label{carmaps_demt_2082}
\end{center}
\end{figure}

@@

{The magnetically open and closed regions of the AWSoM model are associated with Coronal holes (CHs) and the equatorial streamer belt, respectively. The {location of the} open/closed boundaries derived from the respective AWSoM model quite accurately matches {the regions of the DEMT maps, which exhibit the strongest latitudinal gradient of both the electron density and temperature.}}

@@

Latitude-longitude maps of the score $R$ defined by Equation \ref{R} show that \edit{the agreement between the tomographic and synthetic FBEs meets $R \lesssim 0.01$ in more than 90\% of the voxels where tomographic reconstruction is possible (i.e. where FBEs are non zero), and meets $R \lesssim 0.10$ in the rest. This implies that the LDEM found in each voxel accurately predicts the reconstructed FBEs.} \comment{Diego, evaluá que el valor que incluí es correcto. Simplemente fijate en que fracción de los voxels SIN ZDAs $R<0.01$ para CR-2208, yo estimo $90\%$. pra 2082 no hace falta. Comentame antes de enviar.}

. The notable exception is to be found only in the CHs of the rotation {analysed} based on AIA data. A similar results was found in the two existing DEMT works based on data provided by the AIA instrument \citep{nuevo_2015,maccormack_2017}. This point is further discussed below in this section.

@@

\edit{For both rotations, the population of type 0 (small-closed-down) legs population peaks in the core of the streamer belt, around the equator in the case of CR-2082, and towards the southern hemisphere in the case of CR-2208.} This kind of structure was originally found by \citet{huang_2012}, and their existence was shown to be anti-correlated with the solar activity level around the solar minimum between {SCs 23 and 24} by \citet{nuevo_2013}. Later on, \citet{lloveras_2017} showed {that} equatorial down loops in streamers were also to be found in the deep minimum between SCs {23 and 24}. Here, we verify the existence of this type of structure for the two rotations. The relatively smaller population of down legs seen in CR-2208, as compared to CR-2082, is consistent with the aforementioned results by \citet{nuevo_2013}. \edit{Type I (small-closed-up) legs are present at all latitudes in the streamer belt. Their population peaks in the mid-latitudes of both hemispheres for CR-2082, and in the mid-latitudes of the northern hemisphere for CR-2208. Type II (large up) legs are mostly very large trans-equatorial field lines forming the envelope of the streamer belt (the requirement of footpoints located beyond mid-latitudes was included precisely to select this kind of loop). Finally, type III (open) legs populate the high latitude CHs.}

>>>>

{As detailed in Section \ref{intro}, rotations CR-2082 and CR-2208 were selected to carry out DEMT reconstructions, based on data taken by the STEREO/EUVI-B and SDO/AIA instruments, respectively.} The EUVI and AIA data were prepared using the latest processing tools and calibration corrections provided by their teams through the SolarSoft package. \edit{In the case of EUVI data, stray-light contamination is removed by deconvolution of the point spread function (PSF), carefully determined for each detector by \citet{shearer_2012}. In the case of AIA data, we have not yet implemented such a procedure as we were not aware of reliable determinations of their PSF, and we also understand that stray light contamination is expected to be not so important for this instrument. A recent study by \citet{saqri_2020} indicates that the effect is noticeable in DEM analysis of coronal holes (CHs). That is also the case for DEM analysis of EUVI images. Nonetheless, as shown by \citet{lloveras_2017}, due to the temporal and spatial binning of the images used in tomography the effect of stray-light removal in DEMT products turns out to be mild, being $\lesssim 10\%$ for density products and negligible for temperature products. In the future, we will explore the effect of stray-light removal in AIA images on DEMT tomography, which we expect to be smaller than for EUVI images.} For this work, we introduced two improvements in the implementation of the DEMT technique, as described next.

{While in all previous DEMT studies full-disk data was used to perform tomography, in this work we {opt} to only use off-limb data. In this way, the smallest scale and brighter coronal features seen on disk (most typically in the 171\AA\ band) are not included. This has two implications. Firstly, the fast dynamics that typically characterizes those structures is absent from the data. Secondly, only half synodic rotation worth of data is needed to constrain the inversion problem for the whole coronal volume. As a result, {artifacts induced by coronal dynamics} are reduced compared to previous DEMT reconstructions.}

@@

\edit{As the LDEM describes the temperature distribution of the plasma in a specific voxel, it does not deal with different large scale structures, as it may be the case for the DEM describing the plasma along a full LOS. As a result, LDEMs are usually successfully modeled with simpler profiles (such as Gaussian) than those returned by DEM studies constrained by LOS-integrated intensities. Parametric techniques are also used for DEM analysis of narrowband images, such as in the works by \citet{aschwanden_2011,plowman_2013,delzanna_2013}. Other methods applied to DEM analysis of narrow band images include Monte Carlo Markov Chain (MCMC) techniques \citep{schmeltz_2016}, regularized inversion techniques \citep{hannah_2012}, and iterative solvers \citep{pickering_2019,morgan_2019} that use the known TRFs of all filters as a functional base. The latter work in particular, introduces a fast iterative solver named SITES, which can easily be adapted for its use in DEMT. We will explore this in the future.}

@@

\edit{In the case of closed-small field lines, the population of down and up legs becomes comparable for CR-2082, so we classify them into the two complementary classes of legs of type 0 (down) and legs of type I (up). In the case of closed-large field lines, legs of type down are $\lesssim 15\%$ of the population for both rotations. In the case of open field lines, legs of type down are $\lesssim 10\%$ of the population for both rotations. Hence, the requirement of being up for legs of type II and III is included to select the vastly dominating population in each case. On the other hand, the inclusion of latitude limits for the footpoints in the classification of legs from type 0 through III is purposely set to study the streamer belt in progressively outer layers, as well as to separate the field lines of the high latitude CHs (legs of type III). In Section \ref{resu}, the results of both the DEMT and AWSoM models in the four classes of legs are statistically analysed.}

@@

\begin{figure}[h!]
\begin{center}
\includegraphics[width=0.495\textwidth]{figs/map_Ne_CR2082_DEMT-EUVI_behind_H1-L3523_r3d_1025_Rsun.pdf}
\includegraphics[width=0.495\textwidth]{figs/map_Tm_CR2082_DEMT-EUVI_behind_H1-L3523_r3d_1025_Rsun.pdf}
\includegraphics[width=0.495\textwidth]{figs/map_Ne_CR2082_DEMT-EUVI_behind_H1-L3523_r3d_1065_Rsun.pdf}
\includegraphics[width=0.495\textwidth]{figs/map_Tm_CR2082_DEMT-EUVI_behind_H1-L3523_r3d_1065_Rsun.pdf}
\includegraphics[width=0.495\textwidth]{figs/map_Ne_CR2082_DEMT-EUVI_behind_H1-L3523_r3d_1105_Rsun.pdf}
\includegraphics[width=0.495\textwidth]{figs/map_Tm_CR2082_DEMT-EUVI_behind_H1-L3523_r3d_1105_Rsun.pdf}
\caption{Carrington maps of DEMT {products $\sqravgN$} (left panels) and $\Tm$ (right panels) for CR-2082. Top, middle and bottom panels show the results at three heliocentric heights, $1.025$, $1.065$ and $1.105\,\mrsun$ respectively. Black voxels correspond to non-reconstructed regions (see text in Section \ref{demt_res}) and thick-black curves indicate the open/closed boundaries.}
\label{carmaps_demt_2082}
\end{center}
\end{figure}

@@

{The magnetically open and closed regions of the AWSoM model are associated \edit{with CHs} and the equatorial streamer belt, respectively. The {location of the} open/closed boundaries derived from the respective AWSoM model quite accurately matches {the regions of the DEMT maps, which exhibit the strongest latitudinal gradient of both the electron density and temperature.}}

@@

Latitude-longitude maps of the score $R$ defined by Equation \ref{R} show that \edit{the agreement between the tomographic and synthetic FBEs is 5\% or better in more than 90\% of the reconstructed coronal volume (i.e. where FBEs are non zero), and of order 10\% in the rest of the volume. This implies that the LDEM found in each voxel accurately predicts the reconstructed FBEs.}

@@

\edit{For both rotations, the population of type 0 (small-closed-down) legs peaks at the core of the streamer belt, around the equator in the case of CR-2082, and towards the southern hemisphere in the case of CR-2208.} This kind of structure was originally found by \citet{huang_2012}, and their existence was shown to be anti-correlated with the solar activity level around the solar minimum between {SCs 23 and 24} by \citet{nuevo_2013}. Later on, \citet{lloveras_2017} showed {that} equatorial down loops in streamers were also to be found in the deep minimum between SCs {23 and 24}. Here, we verify the existence of this type of structure for the two rotations. The relatively smaller population of down legs seen in CR-2208, as compared to CR-2082, is consistent with the aforementioned results by \citet{nuevo_2013}. \edit{Type I (small-closed-up) legs are present at all latitudes within the streamer belt. Their population peaks in the mid-latitudes of both hemispheres for CR-2082, and in the mid-latitudes of the northern hemisphere for CR-2208. The latitude distribution of legs of type 0 and I in CR-2082 is consistent to the distribution of down and up loops of CR-2081 in the analysis by \citet{nuevo_2013}, which did not place any limits on the latitude location of the analysed structures.}

\edit{Type II (large up) legs are mostly very large trans-equatorial field lines forming the envelope of the streamer belt (the requirement of footpoints located beyond mid-latitudes was included precisely to select this kind of loop). Finally, type III (open) legs populate the high latitude CHs.}
